---
title: "Practical Machine Learning Course Project"
output: html_document
---

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal of this project is to predict the manner in which they did the exercise.

## Load the data
```{r}
library(caret)
trainingDataSet<-read.csv("training.csv")
testingDataSet<-read.csv("testing.csv")
summary(trainingDataSet)
summary(testingDataSet)
dim(trainingDataSet)
dim(testingDataSet)
```

## Clean the data
 We remove columns with missing values from the training and the testing datasets.
 
```{r}
trainingDataSet<-trainingDataSet[,colSums(is.na(trainingDataSet))==0]
dim(trainingDataSet)

testingDataSet<-testingDataSet[,colSums(is.na(testingDataSet))==0]
dim(testingDataSet)
```

## Preprocess the data

```{r}
numericalsIdx <- which(lapply(trainingDataSet, class) %in% "numeric")
preprocessModel <-preProcess(trainingDataSet[,numericalsIdx],method=c('knnImpute', 'center', 'scale'))
pre_trainingDataSet <- predict(preprocessModel, trainingDataSet[,numericalsIdx])
pre_trainingDataSet$classe <- trainingDataSet$classe


```

## Removing the near zero variables
Removing the variables with values near zero, that means that they have not so much meaning  in the predictions

```{r}
 nzv <- nearZeroVar(pre_trainingDataSet,saveMetrics=TRUE)
pre_trainingDataSet <- pre_trainingDataSet[,nzv$nzv==FALSE]


```

## Validation set
We want a 75% observation training dataset to train our model. We will then validate it on the last 70%.
```{r}
set.seed(123)
idxTrain<- createDataPartition(pre_trainingDataSet$classe, p=3/4, list=FALSE)
training<- pre_trainingDataSet[idxTrain, ]
validation <- pre_trainingDataSet[-idxTrain, ]
dim(training) 
dim(validation)
```

## Train Model
We train a model using random forest with a cross validation of 5 folds to avoid overfitting.

```{r}
library(randomForest)
modFitrf <- train(classe ~., method="rf", data=training, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE, importance=TRUE )
modFitrf
```

## Interpretation
Let's plot the importance of each individual variable

```{r}
varImpPlot(modFitrf$finalModel, sort = TRUE, type = 1, pch = 19, col = 1, cex = 0.6, main = "Importance of the Individual Principal Components")
```

## Cross Validation Testing and Out-of-Sample Error Estimate
Let's apply our training model on our testing database, to check its accuracy.

### Accuracy and Estimated out of sample error

```{r}
predValidRF <- predict(modFitrf, validation)
confus <- confusionMatrix(validation$classe, predValidRF)
confus$table

accur <- postResample(validation$classe, predValidRF)
modAccuracy <- accur[[1]]
modAccuracy

out_of_sample_error <- 1 - modAccuracy
out_of_sample_error
```

## Application of this model on the 20 test cases provided
We have already clean the test data base (teData). We delete the "problem id" column as it is useless for our analysis.

```{r}
pred_final <- predict(modFitrf, testingDataSet)
pred_final
```



